---
title: "Exercise Set 1"
format: 
  pdf:
    default-image-extension: png
---


```{r, setup, include=FALSE}
library(targets)
library(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, include = FALSE}
tar_load_everything(strict = F, silent=T)
```

This report contains my answers for exercise set 1.

## Problem 1

### Task a

Drop columns `id` and `partlybad`:

```
dt[, c("id", "partlybad") := NULL]

```

### Task b

Summary of columns `T84.mean`, `UV_A.mean` and `CS.mean`:

```{r}
#| echo: false
print(e1p1_b)
```

### Task c

List of `t84.mean` mean and standard deviation:

```{r}
#| echo: false
print(e1p1_c)
```

### Task d


```{r}
#| echo: false
e1p1_d_path <- normalizePath(e1p1_d)
include_graphics(e1p1_d_path, rel_path = FALSE)
```



### Task e

Scatterplot matrix of `UV_A.mean`,`T84.mean` and `H2O84.mean`:

```{r}
#| echo: false
e1p1_e_path <- normalizePath(e1p1_e)
include_graphics(e1p1_e_path, rel_path = FALSE)
```


### Task f

A dummy model was created by first determining the most common class in column `class4` which was `nonevent`.
Then the probability for an event was calculated by `p_event = nrows events/total nrows`. The ```event``` frequencies
in the data summed up to 225 which ìs the same value as the frequency of `nonevent` values which means that `p_event=0.5`.
These dummy predictions produced a score of `0.36269` on the Kaggle leaderboard.


## Problem 2


### Task a

MSE table for synthetic data:


```{r}
#| echo: false
kable(e1p2_a)
```

To choose the polynomial order when given a combined training and validation set but no test set I would look at the
smallest error value in the table and find out its value in th `Degree` column. I would only be looking at the `Validation` and `CV`
columns as they are the ones available for this task. The resulting table:

```{r}
#| echo: false
kable(e1p2a_min_vals)
```

As the table shows I should train the model on the training set with `Degree` set to `3` and test it on the validation set.


### Task b

Plots of synthetic training set points x and y and fitted polynomials with different values of p:

```{r}
#| echo: false
p2_b_path <- normalizePath(e1p2_b_plots)
include_graphics(p2_b_path, rel_path = FALSE)
```


### Task c

Table of RMSE from fitting different regressors on the real data:


```{r}
#| echo: false
kable(e1p2_c)
```

Answers:

1. The Random Forest (RF) is the best regressor because it produces the smallest values for all cases as shown below:

```{r}
#| echo: false
kable(e1p2c_min_vals)
```


2. Excluding the *Dummy* model, the values of *Train* are smaller than those of *Test*. The same is true for *CV*.
The *RF* *Train* value is substantially lower at *0.5760683* compared to a *Test* value of *1.4510838*.

3.The regressors could be improved by excluding certain predictors from the data. Further analysis should
be done to find out which ones.

## Problem 3

### Task a

1. Training error: The MSE keeps decreasing as flexibility increases. 
2. Test error: A U-shaped curve. Error decreases and then increases again after reaching a minimum point.
3. (squared) bias: In general the more flexibility a model has, the less bias there is because the model is able to follow
the patterns in the data more closely.

4. Variance: In general the variance grows as flexibility increases because with large flexibility even a small change to 
the data can result in a big change in $\hat{f}$.

5.irreducible (or Bayes) error: The irreducible error curve/line shows the threshold of the lowest achievable MSE for any method and 
therefore it is constant.


### Task b

(i) Table of terms for each degree:

```{r}
#| echo: false
kable(e1p3_b1)
```


(ii) Plots:

```{r}
#| echo: false
e1p3_b2_path <- normalizePath(e1p3_b2)
include_graphics(e1p3_b2_path, rel_path = FALSE)
```


(iii) The Irreducible error is constant which makes sense, however the other curves look really similar to each other.
The variance increases a lot when the degree=6. The Total and MSE are equal.


## Problem 5

### Task a

Summary data of fitted models:

```{r}
#| echo: false
kable(e1p5a_dt)
```

The slope terms suggest that when x increases 1 unit, y will increase around 0.5 units in each case.

### Task b

None of the models fit the data very well. The line in d1 looks like it is close to the mean of the observations though.

```{r}
#| echo: false
e1p5b_path <- normalizePath(e1p5b)
include_graphics(e1p5b_path, rel_path = FALSE)
```


### Task c

Non-linearity of the response-predictor relationships applies to d1 and d2. D3 and d4 are linear but each has one point that
is outside the normal range. For d3 this looks like an outlier whereas for d4 it could be a high leverage point.

Plotting residuals is a way to identify the linearity/non-linearity of the data. Below are plots of the residuals vs predictor x for 
each data set. D1 implies linearity because there is little indication of a pattern in the residuals. D2 however has a U-shape which is 
a strong indication of non-linearity. The decreasing line in D3 may be due to the outlier in the data. The values of x in d4 are mostly 
constant and so not much can be said about linearity.

```{r}
#| echo: false
e1p5c_path <- normalizePath(e1p5c)
include_graphics(e1p5c_path, rel_path = FALSE)
```


##  Problem 6

### Task a

```{r}
#| echo: false
print("Bootstrap standard errors:")
print(apply(e1p6a$t, 2, sd))
cat("\n")
print("Original values:")
print(e1p6a$t0)
```

After using bootstrap with `R=5000` re-samples, the `std. error` of the result shows that the intercept `t1` is clearly more unstable than the slope `t2`. This means that for the intercept `t1`, the std. error varies `1.5441767` across re-samples when the slope `t2` only varies `0.1636514`. This is 
also confirmed by the biases which are `t1=0.131211434` and `t2=-0.001837802`. A smaller bias indicates that the original value is a good estimate.

## Task b

The bootstrap algorithm compares the results of a statistic using $n$ re-samples of the original data. In the above exercise it re-samples from `d2.csv`
and fits a linear model for each re-sample. It then measures how accurate the original statistic (intercept and slope in this case) was based on the 
results it gets for the re-sampled data. So the bootstrap is computing the `standard deviation` of the statistic.

## Task c

The probability of an observation to end up in the bootstrap sample is always constant because of replacement. 
So for a population of $n$ sampling $n$ data points the probability of the $j$th observation to not be in the sample is $P=(n-1/n)^n$.
So if $n$ = 10000 then $P = \left(\tfrac{9999}{10000}\right)^{10000} \approx 0.368$.






