---
title: "Exercise Set 2"
format: 
  pdf:
    default-image-extension: png
---


```{r, setup, include=FALSE}
library(targets)
library(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, include = FALSE}
tar_load_everything(strict = F, silent=T)
```

This report contains my answers for exercise set 2.

## Problem 8

### Task a

Coefficients for penguins train data fit.

```{r}
#| echo: false
print(coef(e2p8_a$fit))
```

Accuracy on training and test sets without regularisation:

```{r}
#| echo: false
print("Train accuracy:")
print(e2p8_a$train_acc)
print("Test accuracy:")
print(e2p8_a$test_acc)
```


Plot of penguins train data predictions.


```{r}
#| echo: false
e2p8a_path <- normalizePath(e2p8_a_plot)
include_graphics(e2p8a_path, rel_path = FALSE)
```



## Task b

Coefficients for penguins train data fit.

```{r}
#| echo: false
print(e2p8_b$coef_lasso)
```


Accuracy on training and test sets using lasso with regularisation:

```{r}
#| echo: false
print("Train accuracy:")
print(e2p8_b$train_acc)
print("Test accuracy:")
print(e2p8_b$test_acc)
```

## Task c

R gives warnings because the model is fitting the data too well and because there's no regulation, it will try to push probabilities closer to
0 or 1 forever if the programme didn't stop itself and give the warning.















